### 2024-04-22 paper/paper.qmd
Update paper and datasheet

---
title: "With great growth, comes greater Consumption: Energy consumption of countries around the world"
subtitle: "An analysis of economic growth and Energy consumption globally"
author: 
  - Rayan Awad Alim
thanks: "Code and data are available at: https://github.com/RayanAlim/Energy-Analysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Install Packages if not downloaded:
install_if_missing <- function(package_name) {
  if (!requireNamespace(package_name, quietly = TRUE)) {
    install.packages(package_name)
    suppressMessages(library(package_name, character.only = TRUE))
  }
}

install_if_missing("readxl")
install_if_missing("here")
install_if_missing("dplyr")
install_if_missing("ggplot2")
install_if_missing("kableExtra")
install_if_missing("tidyr")
install_if_missing("caret")
```

```{r}
#| include: false
#| warning: false
#| message: false

# Load 
library(readxl)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(tidyr)
library(caret)

#Load Data 
#data <- read.csv(here("data", "analysis_data", "IDMC_processed.csv"))
```

# Introduction

Economic growth and energy consumption are deeply interlinked processes that influence each other in both direct and indirect ways. Historically, the economic output of a country has been closely tied to its energy usage, with periods of high growth accompanied by increases in energy consumption [@Stern-2006] This relationship is often conceptualized through the lens of the Energy-Economic Growth Nexus which posits that energy is a critical input in economic production and, conversely, economic growth can lead to increased energy consumption due to higher industrial activity, transportation needs, and residential and commercial usage [@SORRELL-2009].

Several theoretical frameworks have been proposed to explain this nexus. The most prominent among them is the Jevons Paradox, which suggests that as technological advancements increase energy efficiency, the rate of consumption of that energy may actually increase as its effective cost decreases [@ALCOTT-2005]. Conversely, the Khazzoom-Brookes postulate extends this paradox to the macroeconomic level, arguing that increased energy efficiency may lead to faster economic growth, which in turn increases total energy demand [@Saunders-1992].

Empirical studies have shown diverse results depending on the economic structure and developmental stage of the country in question. For developed nations, research has indicated a decoupling of energy consumption and economic growth, attributed to advances in energy efficiency and a shift towards service-oriented economies which are less energy-intensive [@OWEN-2006]. In contrast, developing countries tend to exhibit a strong correlation between energy consumption and GDP growth, driven by industrialization and infrastructure expansion which are highly energy-dependent [@Bhattacharyya-2019].

The policy implications of these findings are significant, as they influence national energy strategies and their alignment with economic objectives. The transition towards sustainable energy sources is also a critical factor in this equation, with renewable energy adoption seen as a key element in sustaining long-term economic growth without the environmental degradation associated with fossil fuels [@Wei-2023].

Understanding the specific dynamics of energy consumption and economic growth within the context of a particular country requires a detailed examination of its economic policies, energy resources, and technological advancements. This paper aims to explore these themes within the context of countries globally, providing insights that may help in formulating strategies for sustainable economic and energy development.

# Data {#sec-data}

## Variables

## Dataset

## Plots

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## Data and Model Findings

### First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

### Second discussion point

### Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {A}

## Datasheet

Extract of the questions from @gebru2021datasheets.
Extract of the questions from @gebru2021datasheets.
Datasets used to create this datsetet are: [@citeEnergyConsumption2023] and [@]

**Dataset Abstract**

This datasheet presents a dataset combining annual GDP growth data from the World Bank and OECD National Accounts with per capita primary energy consumption data from the U.S. Energy Information Administration and the Energy Institute's Statistical Review of World Energy. The dataset spans from 1965 to 2022 and includes information for 266 countries, enabling analyses of the relationships between economic growth and energy usage.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was specifically created to examine the relationship between economic growth, measured through GDP growth rates, and energy consumption on a per capita basis across various countries and years. The intent was to analyse the link in economic metrics with energy usage statistics in an effort to understand how economic trends are related to energy consumption pattern.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - There are two datasets merged into one in this dataset. |The first dataset is from the  World Bank national accounts data, and OECD National Accounts data files and the second datset is from U.S. Energy Information Administration (2023), Energy Institute - Statistical Review of World Energy (2023), Population based on various sources (2023) – with major processing by Our World in Data. They were merged by Rayan Awad Alim, they were merged for research purposes and the above outlined motivation.

3. *Who funded the creation of the dat9aset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The collection of the data was funded by World Bank and Organisation for Economic Co-operation and Development (OCED).

4. *Any other comments?*
    - This dataset coule be particularly valuable for policymakers and researchers interested in the implications of economic policies on energy consumption and environmental impact.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance in the dataset represents a country-year pair, with metrics for GDP growth and energy consumption per capita for that particular year and country

2. *How many instances are there in total (of each type, if appropriate)?*
	-     The dataset includes data from 1965 to 2022 for x countries, as well as by region. Each year's data for each country forms a unique instance in the dataset.

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is not a sample it includes all countrries recognized by the UN aqre accounted for.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance includes the following data points; Country name, Year, Annual GDP growth rate (percentage), Energy consumption per capita (kilowatt-hours per person)

5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- There are no explicit labels or target variables as this dataset is designed for exploratory and inferential statistical analyses rather than predictive modeling.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some instances have missing data due to non-reporting countries or years where the data collection was not feasible. Missing data points are noted and are a dropped in the analysis to avoid biased interpretations.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- The dataset primarily represents independent instances without explicit relational data between them

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- N/A

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Some country names werew spelled differenltly, but they were standardized using their ISO3 names.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained with regards to the information it provides. External links to the data sources are maintained for verification and detailed reference but are not necessary for the primary analysis tasks the dataset is intended to do

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- yes, it identifies regions for country clusters, eg africa, north africa, middle east etc

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data for this dataset was sourced from authoritative international databases including the World Bank, OECD National Accounts data files, the U.S. Energy Information Administration (EIA), and the Energy Institute's Statistical Review of World Energy. These sources provide quantifiable, directly observable data compiled from governmental and energy sector reporting.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Not a sample, the entire dataset.

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data compilation and preliminary analysis were conducted by a team of data scientists and economists employed by the respective data-providing organizations. These individuals are experts in their fields and are compensated as full-time employees by their institutions.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The dataset includes data collected annually from 1965 to 2022. Each data point is recorded after the end of the reporting year.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No information provided.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- The data is publicly available in these 2 sites https://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG?end=2022&start=2022&view=map

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- R language [@citeR]

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Only for the analysis this datasheet is accompanying.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No. This dataset will exist in the following github repo which has an MIT license https://github.com/RayanAlim/Energy-Analysis/

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The authors do not intend to maintain this dataset, but it will exist in the following github repo which has an MIT license for others to use  https://github.com/RayanAlim/Energy-Analysis/

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- rayan.alim@mail.utoronto.ca

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- - The authors do not intend to maintain this dataset, but it will exist in the following github repo which has an MIT license for others to use  https://github.com/RayanAlim/Energy-Analysis/.

\newpage
\newpage

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References